---
title: "explore if the parameterization works well for different Continents"
author: "Yunpeng"
date: "1/12/2022"
output: html_document
---
In this script,using Koen's stress function

-for different continents:NorthAmerica, Europe, East Asia(only two sites in Japan)


### r setup 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(GenSA)
source("../R/frost_hardiness.R")
```

### (1) read data
```{r message=FALSE,warning=FALSE}
#load the data uploaded by Koen
df_recent <- readRDS("../data/model_data.rds") %>%
  mutate(
    year = format(date, "%Y")
  ) %>%
  na.omit()

#load the data Beni sent me before:
df_old<-read.csv(file=paste0("../data/","ddf_fluxnet2015_pmodel_with_forcings_stocker19gmd.csv"))
df_old<-df_old %>%
  mutate(date=lubridate::mdy(date),
         year=lubridate::year(date)) %>%
  na.omit(gpp_obs,)
```

### (2) retreive the optimized parameter for the selected sites

```{r echo=FALSE, warning= FALSE}
# set initial value
par <- c(a = 1.5,b = -21.5,T8 = 11.3,t = 5,base_t = -25)

lower=c(0,-100,-50,1, -100)
upper=c(100,0,100,90, 0)

# run model and compare to true values
# returns the RMSE
cost <- function(
  data,
  par
  ) {

  scaling_factor <- data %>%
    # group_by(sitename) %>%
    do({
      scaling_factor <-frost_hardiness(
        .$tmin,
        par
      )

      data.frame(
        sitename = .$sitename,
        date = .$date,
        scaling_factor = scaling_factor
      )
    })

  df <- left_join(data, scaling_factor)

  rmse <- sqrt(
    sum(
      (df$gpp - df$gpp_mod * df$scaling_factor)^2)
    )/nrow(df)

  # This visualizes the process,
  # comment out when running for real
  # plot(df$gpp, type = 'p',ylim=c(0,12))
  # lines(df$gpp_mod, col = "red")
  # lines(df$gpp_mod * df$scaling_factor, col = "blue",cex=1.2)
  # Sys.sleep(0.1)

  return(rmse)
}

```

### optimize for each continents

```{r  echo=FALSE,warning= FALSE}
#first load the continents information:
#load the modis data-->tidy from Beni
df_sites_modis_era <- read_rds(paste0("../data/df_sites_modis_era.csv"))
#add the continent information:
df_sites_modis_era$continent<-rep(NA,nrow(df_sites_modis_era))
pos_NorthAmerica<-c(grep("CA",df_sites_modis_era$sitename),
              grep("US",df_sites_modis_era$sitename))
pos_EastAsia<-c(grep("JP",df_sites_modis_era$sitename))
pos_Europe<-setdiff(c(1:nrow(df_sites_modis_era)),
                    c(pos_NorthAmerica,pos_EastAsia))
df_sites_modis_era$continent[pos_Europe]<-"Europe"
df_sites_modis_era$continent[pos_EastAsia]<-"EastAsia"
df_sites_modis_era$continent[pos_NorthAmerica]<-"NorthAmerica"

#
df_merge<-df_recent %>%
left_join(
    df_sites_modis_era,
    by = "sitename"
  ) 
#main continents
continents<-unique(df_merge$continent)
# ## optimize for each PFT
# library(tictoc)#-->record the parameterization time
# tic("start to parameterize")
# par_continents<-c()
# for(i in 1:length(continents)){
#   df_sel<-df_merge %>%
#     dplyr::filter(continent==continents[i])
# 
#   optim_par <- GenSA::GenSA(
#   par = par,
#   fn = cost,
#   data = df_sel,
#   lower = lower,
#   upper = upper,
#   control = list(max.call=100))$par
# 
#   print(i)
#   par_continents[[i]]<-optim_par
# }
# print("finish parameterization")
# toc()
# 
# names(par_continents)<-continents
# print(par_continents)
# #save the optimized data
# save(par_continents,file = paste0("../data/parameters/","optim_par_run100_koen_continents.rds"))
```

### compare the gpp_obs, ori modelled gpp, and gpp modelled using optimated parameters

print the optim parameters

```{r message=FALSE,warning=FALSE}
load(paste0("../data/parameters/","optim_par_run100_koen_continents.rds"))
#include parameter-->include=FALSE-->Hide output results
print(par_continents)
```

```{r echo=FALSE，message=FALSE,warning=FALSE,include=FALSE}
#--------------------
#(1) get the stress function for each site and merge each sites datasets
#--------------------
# using the optimilized par
df_final<-c()
for (i in 1:length(continents)) {
  df_sel<-df_merge %>%
    dplyr::filter(continent==continents[i])
  
  scaling_factors <- df_sel %>%
  # group_by(sitename, year) %>%
  do({
    scaling_factor <- frost_hardiness(.$tmin,par_continents[[i]])
    data.frame(
      sitename = .$sitename,
      date = .$date,
      scaling_factor_optim = scaling_factor
    )
  })
  df_sel <- left_join(df_sel, scaling_factors)
  
  #merge different sites:
  df_final<-rbind(df_final,df_sel)
}

#-------------------
#(2) calculate the simple stats for continents
#-------------------
#!!first need to merge the modelled gpp from different sources:
df_final$year<-lubridate::year(df_final$date)
df_merge_new<-left_join(df_final,df_old,by = c("sitename", "date", "year")) %>%
  mutate(gpp_obs_recent=gpp,
         gpp_obs_old=gpp_obs,
         gpp_mod_FULL_ori=gpp_mod_FULL,
         gpp_mod_recent_ori=gpp_mod,
         gpp_mod_recent_optim=gpp_mod*scaling_factor_optim,
         gpp=NULL,
         gpp_obs=NULL,
         gpp_mod=NULL)

#
library(sirad)
df_stats<-c()
for (i in 1:length(continents)){
  df_each<-df_merge_new %>%
    dplyr::filter(continent==continents[i])
  #
  ori_model_evas<-modeval(df_each$gpp_mod_FULL_ori,df_each$gpp_obs_old,stat=c("N","pearson","MAE","RMSE","R2","slope","intercept","EF"))
  recent_model_evas<-modeval(df_each$gpp_mod_recent_ori,df_each$gpp_obs_recent,stat=c("N","pearson","MAE","RMSE","R2","slope","intercept","EF"))
  optim_par_model_evas<-modeval(df_each$gpp_mod_recent_optim,df_each$gpp_obs_recent,stat=c("N","pearson","MAE","RMSE","R2","slope","intercept","EF"))
  
  #
  stats_evals<-rbind(ori_model_evas,recent_model_evas,optim_par_model_evas)
  stats_evals
  #merge different sites
  df_stats[[i]]<-stats_evals
}
names(df_stats)<-continents
print(df_stats)
```

### make evaluation plots

(1) For General plots
```{r echo=FALSE,warning=FALSE,message=FALSE,include=FALSE}
#evaluation using Beni's function:
devtools::load_all("D:/Github/rbeni/")
library(rbeni) #-->make the evaluation plot
library(cowplot)
library(grid)
```

```{r echo=FALSE,warning=FALSE,message=FALSE}
#--------------------------
#modelled and observed gpp:scatter plots
#-------------------------
plot_modobs_general<-c()
df_modobs<-c()
for(i in 1:length(continents)){
  #
  df_modobs_each<-df_merge_new %>%
  filter(continent==continents[i]) %>%
  select(sitename,date,continent,gpp_obs_recent,gpp_mod_FULL_ori,gpp_mod_recent_ori,gpp_mod_recent_optim) %>%
  mutate(gpp_obs=gpp_obs_recent,
         gpp_mod_old_ori=gpp_mod_FULL_ori,
         gpp_mod_recent_ori=gpp_mod_recent_ori,
         gpp_mod_recent_optim=gpp_mod_recent_optim) %>%
  mutate(gpp_obs_recent=NULL,
         gpp_mod_FULL_ori=NULL)
#
df_modobs<-rbind(df_modobs,df_modobs_each)

#scatter plots to compare the model and observation gpp
gpp_modobs_comp1<-df_modobs_each %>%
  analyse_modobs2("gpp_mod_old_ori", "gpp_obs", type = "heat")
gpp_modobs_comp2<-df_modobs_each %>%
  analyse_modobs2("gpp_mod_recent_ori", "gpp_obs", type = "heat")
gpp_modobs_comp3<-df_modobs_each %>%
  analyse_modobs2("gpp_mod_recent_optim", "gpp_obs", type = "heat")
# add the site-name:
gpp_modobs_comp1$gg<-gpp_modobs_comp1$gg+
  annotate(geom="text",x=15,y=0,label=continents[i])
gpp_modobs_comp2$gg<-gpp_modobs_comp2$gg+
  annotate(geom="text",x=15,y=0,label=continents[i])
gpp_modobs_comp3$gg<-gpp_modobs_comp3$gg+
  annotate(geom="text",x=15,y=0,label=continents[i])

#merge two plots
evaulation_merge_plot<-plot_grid(gpp_modobs_comp1$gg,
    gpp_modobs_comp2$gg,gpp_modobs_comp3$gg,
    widths=15,heights=4,
labels = "auto",ncol =3,nrow = 1,label_size = 12,align = "hv")
# plot(evaulation_merge_plot)

#put all the plots together:
plot_modobs_general[[i]]<-evaulation_merge_plot
}
names(plot_modobs_general)<-continents

#print the plot
plot_modobs_general

```

(2) For Seasonality

a. Seasonal course for different continents:
```{r echo=FALSE,warning=FALSE,message=FALSE}

#plotting:
  #Seasonal course by climate zone:
  season_plot<-df_modobs %>%
  mutate(doy = lubridate::yday(date)) %>%
  group_by(continent, doy) %>%
  summarise(obs = mean(gpp_obs, na.rm = TRUE),
            mod_old_ori=mean(gpp_mod_old_ori, na.rm = TRUE),
            mod_recent_ori=mean(gpp_mod_recent_ori, na.rm = TRUE),
            mod_recent_optim=mean(gpp_mod_recent_optim,na.rm = TRUE)) %>%
  pivot_longer(c(obs,mod_old_ori,mod_recent_ori,mod_recent_optim), names_to = "Source", values_to = "gpp") %>%
  ggplot(aes(doy, gpp, color = Source)) +
  geom_line() +
  scale_color_manual(values = c("mod_old_ori" = "red","mod_recent_ori"="steelblue2",
                                "mod_recent_optim" = "orange", "obs" = "black"),
                     labels = c("Old P-model","Recent Ori P-model", "Recent Optim P-model","Obs.")) +
  labs(y = expression( paste("GPP (g C m"^-2, " d"^-1, ")" ) ),
       x = "Day of year") +
  annotate(geom="text",x=200,y=2,label="")+
  facet_wrap(~continent)
  
#print the plot
season_plot
```

b. Seasonal course for each sites in different continents:
 
 - for each site in DBF
```{r echo=FALSE,warning=FALSE,message=FALSE}
#Seasonal course for each sites in different continents:
# - for Europe
df_modobs %>%
  filter(continent=="Europe") %>%
  mutate(doy = lubridate::yday(date)) %>%
  group_by(sitename, doy) %>%
  summarise(obs = mean(gpp_obs, na.rm = TRUE),
            mod_old_ori=mean(gpp_mod_old_ori, na.rm = TRUE),
            mod_recent_ori=mean(gpp_mod_recent_ori, na.rm = TRUE),
            mod_recent_optim=mean(gpp_mod_recent_optim,na.rm = TRUE)) %>%
  pivot_longer(c(obs,mod_old_ori,mod_recent_ori,mod_recent_optim), names_to = "Source", values_to = "gpp") %>%
  ggplot(aes(doy, gpp, color = Source)) +
  geom_line() +
  scale_color_manual(values = c("mod_old_ori" = "red","mod_recent_ori"="steelblue2",
                                "mod_recent_optim" = "orange", "obs" = "black"),
                     labels = c("Old P-model","Recent Ori P-model", "Recent Optim P-model","Obs.")) +
  labs(y = expression( paste("GPP (g C m"^-2, " d"^-1, ")" ) ),
       x = "Day of year") +
  facet_wrap(~sitename)

```

- for each site in North America
```{r echo=FALSE,warning=FALSE,message=FALSE}
#Seasonal course for each sites in different continents:
# - for NorthAmerica
df_modobs %>%
  filter(continent=="NorthAmerica") %>%
  mutate(doy = lubridate::yday(date)) %>%
  group_by(sitename, doy) %>%
  summarise(obs = mean(gpp_obs, na.rm = TRUE),
            mod_old_ori=mean(gpp_mod_old_ori, na.rm = TRUE),
            mod_recent_ori=mean(gpp_mod_recent_ori, na.rm = TRUE),
            mod_recent_optim=mean(gpp_mod_recent_optim,na.rm = TRUE)) %>%
  pivot_longer(c(obs,mod_old_ori,mod_recent_ori,mod_recent_optim), names_to = "Source", values_to = "gpp") %>%
  ggplot(aes(doy, gpp, color = Source)) +
  geom_line() +
  scale_color_manual(values = c("mod_old_ori" = "red","mod_recent_ori"="steelblue2",
                                "mod_recent_optim" = "orange", "obs" = "black"),
                     labels = c("Old P-model","Recent Ori P-model", "Recent Optim P-model","Obs.")) +
  labs(y = expression( paste("GPP (g C m"^-2, " d"^-1, ")" ) ),
       x = "Day of year") +
  facet_wrap(~sitename)

```


- for each site in East Asia
```{r echo=FALSE,warning=FALSE,message=FALSE}
#Seasonal course for each sites in different continents:
# - for East Asia
df_modobs %>%
  filter(continent=="EastAsia") %>%
  mutate(doy = lubridate::yday(date)) %>%
  group_by(sitename, doy) %>%
  summarise(obs = mean(gpp_obs, na.rm = TRUE),
            mod_old_ori=mean(gpp_mod_old_ori, na.rm = TRUE),
            mod_recent_ori=mean(gpp_mod_recent_ori, na.rm = TRUE),
            mod_recent_optim=mean(gpp_mod_recent_optim,na.rm = TRUE)) %>%
  pivot_longer(c(obs,mod_old_ori,mod_recent_ori,mod_recent_optim), names_to = "Source", values_to = "gpp") %>%
  ggplot(aes(doy, gpp, color = Source)) +
  geom_line() +
  scale_color_manual(values = c("mod_old_ori" = "red","mod_recent_ori"="steelblue2",
                                "mod_recent_optim" = "orange", "obs" = "black"),
                     labels = c("Old P-model","Recent Ori P-model", "Recent Optim P-model","Obs.")) +
  labs(y = expression( paste("GPP (g C m"^-2, " d"^-1, ")" ) ),
       x = "Day of year") +
  facet_wrap(~sitename)

```

