---
title: "explore if the parameterization works well for different PFTs"
author: "Yunpeng"
date: "17/12/2021"
output: html_document
---
In this script,using updated Beni's stress function(I also embedded the day length into function)

**here test MF**

-for different PFTs:


### r setup 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(GenSA)
source("../R/model_hardening_byBeni.R")
source("../R/model_hardening_byBeni_addDL.R")
```

### (1) read data
```{r message=FALSE,warning=FALSE}
#load the data uploaded by Koen
df_recent <- readRDS("../data/model_data.rds") %>%
  mutate(
    year = format(date, "%Y")
  ) %>%
  na.omit()

#load the data Beni sent me before:
df_old<-read.csv(file=paste0("../data/","ddf_fluxnet2015_pmodel_with_forcings_stocker19gmd.csv"))
df_old<-df_old %>%
  mutate(date=lubridate::mdy(date),
         year=lubridate::year(date)) %>%
  na.omit(gpp_obs,)
```

add day length
```{r message=FALSE,warning=FALSE}
#calculate the daylength for each sites
library(geosphere) #calculate the day length
library(sirad)  #calculate the day of the year
library(lubridate)
library(dplyr)
#load the lat and long of the example site:
df_sites_modis_era <-read_rds(paste0("../data/df_sites_modis_era.csv"))
#-------
#adding DL for df_recent datasets
#-------
sites<-unique(df_recent$sitename)
df_recent_temp<-c()
for(i in 1:length(sites)){
  df.temp<-df_recent%>%filter(sitename==sites[i])
  siteinfo<-df_sites_modis_era %>% filter(sitename==sites[i])
  #
  df.temp$doy<-dayOfYear(df.temp$date)
  df.temp$DL<-daylength(siteinfo$lat,df.temp$doy)
  df_recent_temp<-rbind(df_recent_temp,df.temp)
}
#
df_recent<-df_recent_temp

```

### (2) retreive the optimized parameter for the selected sites

```{r echo=FALSE, warning= FALSE}
# set initial value
par_old<- c("a" = 0, "b" = 0.5, "c" = 50, "d" = 0.1, "e" = 1)
par_new1 <- c("a" = 0, "b" = 0.5, "c" = 50, "d" = 0.1, "e" = 1)
par_new2 <- c("a" = 0, "b" = 0.5, "c" = 50, "d" = 0.1, "e" = 1)
par_new3 <- c("a" = 0, "b" = 0.5, "c" = 50, "d" = 0.1, "e" = 1,"f"=6)

#set the parameter variation ranges
lower_old=c(-50,0,0,0,0)
upper_old=c(50,20,100,20,10)
#
lower_new1=c(-50,0,0,0,0)
upper_new1=c(50,20,500,20,10)
#
lower_new2=c(-50,0,0,0,0)
upper_new2=c(50,20,500,20,10)
#
lower_new3=c(-50,0,0,0,0,0)
upper_new3=c(50,20,500,20,10,24)

# run model and compare to true values
# returns the RMSE
# source the cost functions
source(paste0("../R/cost_fun.R"))
```

### optimize for each PFTs

```{r  echo=FALSE,warning= FALSE}
#first load the PFTs information:
#load the modis data-->tidy from Beni
df_sites_modis_era <- read_rds(paste0("../data/df_sites_modis_era.csv"))
#
df_merge<-df_recent %>%
left_join(
    df_sites_modis_era,
    by = "sitename"
  ) 
#main PFTs
# PFTs<-unique(df_merge$classid)
#first target MF:
PFTs<-"MF"
### optimize for each PFT
library(tictoc)#-->record the parameterization time
# tic("start to parameterize")
# par_PFTs<-c()
# for(i in 1:length(PFTs)){
df_sel<-df_merge %>%
    dplyr::filter(classid==PFTs)

##1) for old model strcture
tic("start to parameterize")
optim_par_old <- GenSA::GenSA(
  par = par_old,
  fn = cost_old,
  data = df_sel,
  lower = lower_old,
  upper = upper_old,
  control = list(max.call=100))$par

  # print(i)
  # par_PFTs[[i]]<-optim_par
# }
print("finish parameterization")
toc()
# names(par_PFTs)<-PFTs
# print(par_PFTs)
#save the optimized data
save(optim_par_old,file = paste0("../data/parameters/addNL_test/","optim_par_oldmod_run100_beni_addNL_MF.rds"))

##2) for new model strcture--new model1
tic("start to parameterize")
optim_par_new1 <- GenSA::GenSA(
  par = par_new1,
  fn = cost_new1,
  data = df_sel,
  lower = lower_new1,
  upper = upper_new1,
  control = list(max.call=100))$par

print("finish parameterization")
toc()
#save the optimized data
save(optim_par_new1,file = paste0("../data/parameters/addNL_test/","optim_par_newmod1_run100_beni_addNL_MF.rds"))

##3) for new model strcture--new model2
tic("start to parameterize")
optim_par_new2 <- GenSA::GenSA(
  par = par_new2,
  fn = cost_new2,
  data = df_sel,
  lower = lower_new2,
  upper = upper_new2,
  control = list(max.call=100))$par

print("finish parameterization")
toc()
#save the optimized data
save(optim_par_new2,file = paste0("../data/parameters/addNL_test/","optim_par_newmod2_run100_beni_addNL_MF.rds"))

##4) for new model strcture--new model3
tic("start to parameterize")
optim_par_new3 <- GenSA::GenSA(
  par = par_new3,
  fn = cost_new3,
  data = df_sel,
  lower = lower_new3,
  upper = upper_new3,
  control = list(max.call=100))$par

print("finish parameterization")
toc()
#save the optimized data
save(optim_par_new3,file = paste0("../data/parameters/addNL_test/","optim_par_newmod3_run100_beni_addNL_MF.rds"))

```

### compare the gpp_obs, ori modelled gpp, and gpp modelled using optimated parameters

print the optim parameters

```{r message=FALSE,warning=FALSE}
load(paste0("../data/parameters/addNL_test/","optim_par_oldmod_run100_beni_addNL_MF.rds"))
load(paste0("../data/parameters/addNL_test/","optim_par_newmod1_run100_beni_addNL_MF.rds"))
load(paste0("../data/parameters/addNL_test/","optim_par_newmod2_run100_beni_addNL_MF.rds"))
load(paste0("../data/parameters/addNL_test/","optim_par_newmod3_run100_beni_addNL_MF.rds"))

#include parameter-->include=FALSE-->Hide output results
print(optim_par_old)
print(optim_par_new1)
print(optim_par_new2)
print(optim_par_new3)
```

```{r echo=FALSE，message=FALSE,warning=FALSE,include=FALSE}
#--------------------
#(1) get the stress function value for each site and merge each sites datasets
#--------------------
# using the optimilized par
df_final<-c()
# for (i in 1:length(PFTs)) {
  df_sel<-df_merge %>%
    dplyr::filter(classid==PFTs)
  
##stress factors for differenet model structure
  scaling_factors <- df_sel %>%
  # group_by(sitename, year) %>%
  do({
    scaling_factor_old <- model_hardening(.,optim_par_old)
    scaling_factor_new1 <- model_hardening_new1(.,optim_par_new1)
    scaling_factor_new2 <- model_hardening_new2(.,optim_par_new2)
    scaling_factor_new3 <- model_hardening_new3(.,optim_par_new3)
    data.frame(
      sitename = .$sitename,
      date = .$date,
      scaling_factor_optim_oldmod = scaling_factor_old,
      scaling_factor_optim_newmod1 = scaling_factor_new1,
      scaling_factor_optim_newmod2 = scaling_factor_new2,
      scaling_factor_optim_newmod3 = scaling_factor_new3
    )
  })
  df_sel <- left_join(df_sel, scaling_factors)
  
  #merge different sites:
  df_final<-df_sel
# }

#-------------------
#(2) calculate the simple stats for PFTs
#-------------------
#!!first need to merge the modelled gpp from different sources:
df_final$year<-lubridate::year(df_final$date)
df_merge_new<-left_join(df_final,df_old,by = c("sitename", "date", "year")) %>%
  mutate(gpp_obs_recent=gpp,
         gpp_obs_old=gpp_obs,
         gpp_mod_FULL_ori=gpp_mod_FULL,
         gpp_mod_recent_ori=gpp_mod,
         gpp_mod_recent_optim_oldmod=gpp_mod*scaling_factor_optim_oldmod,
         gpp_mod_recent_optim_newmod1=gpp_mod*scaling_factor_optim_newmod1,
         gpp_mod_recent_optim_newmod2=gpp_mod*scaling_factor_optim_newmod2,
         gpp_mod_recent_optim_newmod3=gpp_mod*scaling_factor_optim_newmod3,
         gpp=NULL,
         gpp_obs=NULL,
         gpp_mod=NULL)

#
library(sirad)
df_stats<-c()
# for (i in 1:length(PFTs)){
  df_each<-df_merge_new %>%
    dplyr::filter(classid==PFTs)  #PFTs[i]-->PFTs
  #
  ori_model_evas<-modeval(df_each$gpp_mod_FULL_ori,df_each$gpp_obs_old,stat=c("N","pearson","MAE","RMSE","R2","slope","intercept","EF"))
  recent_model_evas<-modeval(df_each$gpp_mod_recent_ori,df_each$gpp_obs_recent,stat=c("N","pearson","MAE","RMSE","R2","slope","intercept","EF"))
  recent_model_optim_oldmod_evas<-modeval(df_each$gpp_mod_recent_optim_oldmod,df_each$gpp_obs_recent,stat=c("N","pearson","MAE","RMSE","R2","slope","intercept","EF"))
  recent_model_optim_newmod1_evas<-modeval(df_each$gpp_mod_recent_optim_newmod1,df_each$gpp_obs_recent,stat=c("N","pearson","MAE","RMSE","R2","slope","intercept","EF"))
  recent_model_optim_newmod2_evas<-modeval(df_each$gpp_mod_recent_optim_newmod2,df_each$gpp_obs_recent,stat=c("N","pearson","MAE","RMSE","R2","slope","intercept","EF"))
  recent_model_optim_newmod3_evas<-modeval(df_each$gpp_mod_recent_optim_newmod3,df_each$gpp_obs_recent,stat=c("N","pearson","MAE","RMSE","R2","slope","intercept","EF"))
  #
  stats_evals<-rbind(ori_model_evas,recent_model_evas,recent_model_optim_oldmod_evas,
                     recent_model_optim_newmod1_evas,recent_model_optim_newmod2_evas,
                     recent_model_optim_newmod3_evas)
  # stats_evals
  #merge different sites
df_stats<-stats_evals  #df_stats[[i]]-->df_stats
# }
# names(df_stats)<-PFTs
print(df_stats)
```

### make evaluation plots

(1) For General plots
```{r echo=FALSE,warning=FALSE,message=FALSE,include=FALSE}
#evaluation using Beni's function:
devtools::load_all("D:/Github/rbeni/")
library(rbeni) #-->make the evaluation plot
library(cowplot)
library(grid)
```

```{r echo=FALSE,warning=FALSE,message=FALSE}
#--------------------------
#modelled and observed gpp:scatter plots
#-------------------------
plot_modobs_general<-c()
df_modobs<-c()
for(i in 1:length(PFTs)){
  #
  df_modobs_each<-df_merge_new %>%
  filter(classid==PFTs[i]) %>%
  select(sitename,date,classid,gpp_obs_recent,gpp_mod_FULL_ori,gpp_mod_recent_ori,
         gpp_mod_recent_optim_oldmod,gpp_mod_recent_optim_newmod1,
         gpp_mod_recent_optim_newmod2,gpp_mod_recent_optim_newmod3) %>%
  mutate(gpp_obs=gpp_obs_recent,
         gpp_mod_old_ori=gpp_mod_FULL_ori,
         gpp_mod_recent_ori=gpp_mod_recent_ori,
         gpp_mod_recent_optim_oldmod=gpp_mod_recent_optim_oldmod,
         gpp_mod_recent_optim_newmod1=gpp_mod_recent_optim_newmod1,
         gpp_mod_recent_optim_newmod2=gpp_mod_recent_optim_newmod2,
         gpp_mod_recent_optim_newmod3=gpp_mod_recent_optim_newmod3
         ) %>%
  mutate(gpp_obs_recent=NULL,
         gpp_mod_FULL_ori=NULL)
#
df_modobs<-rbind(df_modobs,df_modobs_each)

#scatter plots to compare the model and observation gpp
gpp_modobs_comp_old_ori<-df_modobs_each %>%
  analyse_modobs2("gpp_mod_old_ori", "gpp_obs", type = "heat")
gpp_modobs_comp_recent_ori<-df_modobs_each %>%
  analyse_modobs2("gpp_mod_recent_ori", "gpp_obs", type = "heat")
gpp_modobs_comp_recent_optim_oldmod<-df_modobs_each %>%
  analyse_modobs2("gpp_mod_recent_optim_oldmod", "gpp_obs", type = "heat")
gpp_modobs_comp_recent_optim_newmod1<-df_modobs_each %>%
  analyse_modobs2("gpp_mod_recent_optim_newmod1", "gpp_obs", type = "heat")
gpp_modobs_comp_recent_optim_newmod2<-df_modobs_each %>%
  analyse_modobs2("gpp_mod_recent_optim_newmod2", "gpp_obs", type = "heat")
gpp_modobs_comp_recent_optim_newmod3<-df_modobs_each %>%
  analyse_modobs2("gpp_mod_recent_optim_newmod3", "gpp_obs", type = "heat")


# add the site-name:
gpp_modobs_comp_old_ori$gg<-gpp_modobs_comp_old_ori$gg+
  annotate(geom="text",x=15,y=0,label=PFTs[i])
gpp_modobs_comp_recent_ori$gg<-gpp_modobs_comp_recent_ori$gg+
  annotate(geom="text",x=15,y=0,label=PFTs[i])
gpp_modobs_comp_recent_optim_oldmod$gg<-gpp_modobs_comp_recent_optim_oldmod$gg+
  annotate(geom="text",x=15,y=0,label=PFTs[i])
gpp_modobs_comp_recent_optim_newmod1$gg<-gpp_modobs_comp_recent_optim_newmod1$gg+
  annotate(geom="text",x=15,y=0,label=PFTs[i])
gpp_modobs_comp_recent_optim_newmod2$gg<-gpp_modobs_comp_recent_optim_newmod2$gg+
  annotate(geom="text",x=15,y=0,label=PFTs[i])
gpp_modobs_comp_recent_optim_newmod3$gg<-gpp_modobs_comp_recent_optim_newmod3$gg+
  annotate(geom="text",x=15,y=0,label=PFTs[i])

#merge two plots
evaulation_merge_plot<-plot_grid(gpp_modobs_comp_old_ori$gg,gpp_modobs_comp_recent_ori$gg,gpp_modobs_comp_recent_optim_oldmod$gg,gpp_modobs_comp_recent_optim_newmod1$gg,gpp_modobs_comp_recent_optim_newmod2$gg,gpp_modobs_comp_recent_optim_newmod3$gg,
    widths=15,heights=10,
labels = "auto",ncol =2,nrow = 3,label_size = 12,align = "hv")
# plot(evaulation_merge_plot)

#put all the plots together:
plot_modobs_general[[i]]<-evaulation_merge_plot
}
names(plot_modobs_general)<-PFTs

#print the plot
plot_modobs_general
```

(2) For Seasonality

a. Seasonal course for different PFTs:

```{r echo=FALSE,warning=FALSE,message=FALSE}

#plotting:
  season_plot<-df_modobs %>%
  mutate(doy = lubridate::yday(date)) %>%
  group_by(classid, doy) %>%
  summarise(obs = mean(gpp_obs, na.rm = TRUE),
            mod_old_ori=mean(gpp_mod_old_ori, na.rm = TRUE),
            mod_recent_ori=mean(gpp_mod_recent_ori, na.rm = TRUE),
            mod_recent_optim_oldmod=mean(gpp_mod_recent_optim_oldmod,na.rm = TRUE),
            mod_recent_optim_newmod1=mean(gpp_mod_recent_optim_newmod1,na.rm = TRUE),
            mod_recent_optim_newmod2=mean(gpp_mod_recent_optim_newmod2,na.rm = TRUE),
            mod_recent_optim_newmod3=mean(gpp_mod_recent_optim_newmod3,na.rm = TRUE)) %>%
  pivot_longer(c(obs,mod_old_ori,mod_recent_ori,mod_recent_optim_oldmod,
            mod_recent_optim_newmod1,mod_recent_optim_newmod2,mod_recent_optim_newmod3), names_to = "Source", values_to = "gpp") %>%
  ggplot(aes(doy, gpp, color = Source)) +
  geom_line() +
  scale_color_manual(values = c("mod_old_ori" = "green","mod_recent_ori"="steelblue2","mod_recent_optim_oldmod" = "blue","mod_recent_optim_newmod1" = "tomato","mod_recent_optim_newmod2" = "red3","mod_recent_optim_newmod3" = "orange","obs" = "black"),
  labels = c("Old P-model","Recent Ori P-model", "Recent Optim old P-model",
             "Recent Optim new1 P-model","Recent Optim new2 P-model","Recent Optim new3 P-model","Obs.")) +
  labs(y = expression( paste("GPP (g C m"^-2, " d"^-1, ")" ) ),
       x = "Day of year") +
  annotate(geom="text",x=200,y=2,label="")+
  facet_wrap(~classid)
  
#print the plot
season_plot
```

b. Seasonal course for each sites in different PFTs:
 
 - for each site in MF
```{r echo=FALSE,warning=FALSE,message=FALSE}
#Seasonal course for each sites in different PFTs:
# - for MF
df_modobs %>%
  filter(classid=="MF") %>%
  mutate(doy = lubridate::yday(date)) %>%
  group_by(sitename, doy) %>%
  summarise(obs = mean(gpp_obs, na.rm = TRUE),
            mod_old_ori=mean(gpp_mod_old_ori, na.rm = TRUE),
            mod_recent_ori=mean(gpp_mod_recent_ori, na.rm = TRUE),
            mod_recent_optim_oldmod=mean(gpp_mod_recent_optim_oldmod,na.rm = TRUE),
            mod_recent_optim_newmod1=mean(gpp_mod_recent_optim_newmod1,na.rm = TRUE),
            mod_recent_optim_newmod2=mean(gpp_mod_recent_optim_newmod2,na.rm = TRUE),
            mod_recent_optim_newmod3=mean(gpp_mod_recent_optim_newmod3,na.rm = TRUE)) %>%
  pivot_longer(c(obs,mod_old_ori,mod_recent_ori,mod_recent_optim_oldmod,
            mod_recent_optim_newmod1,mod_recent_optim_newmod2,mod_recent_optim_newmod3), names_to = "Source", values_to = "gpp") %>%
  ggplot(aes(doy, gpp, color = Source)) +
  geom_line() +
  scale_color_manual(values = c("mod_old_ori" = "green","mod_recent_ori"="steelblue2","mod_recent_optim_oldmod" = "blue","mod_recent_optim_newmod1" = "tomato","mod_recent_optim_newmod2" = "red3","mod_recent_optim_newmod3" = "orange","obs" = "black"),
  labels = c("Old P-model","Recent Ori P-model", "Recent Optim old P-model",
             "Recent Optim new1 P-model","Recent Optim new2 P-model","Recent Optim new3 P-model","Obs.")) +
  labs(y = expression( paste("GPP (g C m"^-2, " d"^-1, ")" ) ),
       x = "Day of year") +
  annotate(geom="text",x=200,y=2,label="")+
  facet_wrap(~sitename)

```
